---
title: "Data Upload MD1"
author: "Julie Norman"
date: '2022-04-11'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

```{r}
library(tidyverse)
library(dplyr)
library(readxl)

```

## Data Upload

### 50 Poorest Counties

https://www.census.gov/data/datasets/2013/demo/saipe/2013-state-and-county.html

```{r}
PoorCounties2013_df <- read_csv("PoorCounties2013_Pvrt.csv")

colsA <- c("State FIPS Code", "County FIPS Code", "Postal Code", "Name")

PoorCounties2013_df <- mutate_at(PoorCounties2013_df, colsA, factor) # character to factor

PoorCounties2013_df
```

### Common Causes of Death

```{r}
COD2013_df <- read_csv("Compressed Mortality, 1999-2016.csv")

COD2013_df <- COD2013_df[ , -which(names(COD2013_df) %in% c("Notes", "Crude Rate"))] #remove empty/ unneeded columns

colsB <- c("State", "State Code", "County", "County Code", "Cause of death", "Cause of death Code")

COD2013_df <- mutate_at(COD2013_df, colsB, factor) # character to factor

COD2013_df$'Postal Code' <- as.factor(state.abb[match(COD2013_df$State, state.name)]) # add state abriev

COD2013_df
```


### Combine Dataframes

```{r}
StateCountyFips_df <- read_excel("StateCountyFips.xlsx")

COD2013_df <- COD2013_df %>% fuzzy_inner_join(StateCountyFips_df,  by = c("County" = "Name"), match_fun = str_detect)
```

```{r}
combo_df <- merge(PoorCounties2013_df, COD2013_df, by = "County FIPS Code")
```


### Find aggregate

```{r}
group_sum <- setNames(aggregate(combo_df$Deaths, list(combo_df$`Cause of death`), sum), c("Cause of Death", "Counts"))
group_sum[order(-group_sum$Counts),] # reorder
```

* Bronchus or lung, unspecified - Malignant neoplasms C34.9
* Acute myocardial infarction, unspecified I21.9
* Atherosclerotic cardiovascular disease, so described I25.0
* Chronic obstructive pulmonary disease, unspecified J44.9
* Atherosclerotic heart disease I25.1
* Hypertensive heart disease without (congestive) heart failure I11
* Breast, unspecified - Malignant neoplasms C50.9
* Cardiac arrest, unspecified I46.9



Top 10 Causes of Death

* Septicaemia, unspecified A41.9
* Colon, unspecified - Malignant neoplasms C18.9
* Bronchus or lung, unspecified - Malignant neoplasms C34.9
* Breast, unspecified - Malignant neoplasms C50.9
* Hypertensive heart disease without (congestive) heart failure I11
* Acute myocardial infarction, unspecified I21.9
* Atherosclerotic cardiovascular disease, so described I25.0
* Atherosclerotic heart disease I25.1
* Cardiac arrest, unspecified I46.9
* Chronic obstructive pulmonary disease, unspecified J44.9


```{r}
group_sum10 <- group_sum[1:10,]

sum(group_sum10$Counts) /sum(group_sum$Counts)

```


## Black popultaion 

U.S. Census Bureau, 2010 American Community Survey
Source: U.S. Census Bureau, 2011 American Community Survey
Source: U.S. Census Bureau, 2012 American Community Survey
Source: U.S. Census Bureau, 2013 American Community Survey
Source: U.S. Census Bureau, 2014 American Community Survey

Source: U.S. Census Bureau, 2011-2015 American Community Survey 5-Year Estimates

U.S. Census Bureau, 2016 American Community Survey 1-Year Estimates
Source: U.S. Census Bureau, 2017 American Community Survey 1-Year Estimates

```{r}
year <- c("2010", "2011", "2012", "2013", "2014", "2015", "2016", "2017", "2018", "2019", "2010", "2011", "2012", "2013", "2014", "2015", "2016", "2017", "2018", "2019")
black_pop <- c(464411, 463364, 462182, 463928, 468231, 459748, 464516,  458367, 456714, 467468, 1109263, 1115359, 1123620, 1128280, 1132556, 1116457, 1134958, 1134333, 1135599, 1130608)
state <- c( "AK", "AK", "AK", "AK", "AK", "AK", "AK", "AK", "AK", "AK", "MS", "MS", "MS", "MS", "MS", "MS", "MS", "MS", "MS", "MS")

year2 <- c("2010", "2011", "2012", "2013", "2014", "2015", "2016", "2017", "2018", "2019", )
BlackAK <- c(464411, 463364, 462182, 463928, 468231, 459748, 464516,  458367, 456714, 467468)
BlackMS <- c(1109263, 1115359, 1123620, 1128280, 1132556, 1116457, 1134958, 1134333, 1135599, 1130608)

lst <- list(year = year,
             black_pop = black_pop,
             state = state)

black_df <- as.data.frame(lst)
black_df$year <- as.numeric(black_df$)

ggplot(black_df, aes(x = year, y = black_pop, fill = state)) + 
  geom_line() 



```





